# Meta information
There should be here the libraries, dataset, odk and sampling frame and any other thing.
Note that some variables were NA'ed (age, etc.) In addition, it seems the weights are already added, and both loops where combined already.

```{r, warning=FALSE, message=FALSE }
## Importing dataset and questionnaire 
library(magrittr)
library(dplyr)
library(readxl)
library(hypegrammaR)

main_dataset <- read.csv("inputs/UKR2007_MSNA20_HH_dataset_main_rcop.csv", na.strings = "")
loop_dataset <- read.csv("inputs/UKR2007_MSNA20_HH_dataset_loop_rcop.csv", na.strings = "")

main_dataset <- main_dataset %>% select_if(~ !(all(is.na(.x)) | all(. == "")))

questions <- read_xlsx("inputs/UKR2007_MSNA20_HH_Questionnaire_24JUL2020.xlsx",sheet="survey")
choices <- read_xlsx("inputs/UKR2007_MSNA20_HH_Questionnaire_24JUL2020.xlsx",sheet="choices")

```

## Creating the questionnaire object

```{r, tidy=FALSE, warning=FALSE, results='hide', message=FALSE}


questionnaire <- load_questionnaire(data = main_dataset,
                                    questions = questions,
                                    choices = choices,
                                    choices.label.column.to.use = "label::English")

```

# Downloading data from kobo server 

```{r include=FALSE}
library(httr)
library(readr) 
library(tidyverse) 
library(jsonlite)
```

## Defining the needed functions
```{r}
kobohr_getforms_csv <-function(url,u,pw){
  rawdata<-GET(url,authenticate(u,pw),progress())
  d_content_csv <-read_csv(content(rawdata,"raw",encoding = "UTF-8"))
}

download_data <-function(url,u,pw){
  
  rawdata<- GET(url,authenticate(u,pw),progress())
  d_content <- read_csv(content(rawdata,"raw",encoding = "UTF-8"),na = c("n/a","")) %>% setNames(gsub("/",".",names(.))) 
  
  d_content

}

create_export <-function(type,lang,fields_from_all_versions,hierarchy_in_labels,group_sep,asset_uid,kobo_user,Kobo_pw){
  api_url_export<-paste0(kobo_server_url,"exports/")
  api_url_asset<-paste0(kobo_server_url,"assets/",asset_uid,"/")
  api_url_export_asset<-paste0(kobo_server_url,"exports/",asset_uid,"/")
  #
  d<-list(source=api_url_asset,
          type=type,
          lang=lang,
          fields_from_all_versions=fields_from_all_versions,
          hierarchy_in_labels=hierarchy_in_labels,
          group_sep=group_sep)
  #fetch data
  result<-httr::POST(url=api_url_export,
                      body=d,
                      authenticate(kobo_user,Kobo_pw),
                      progress()
  )
  
  print(paste0("status code:",result$status_code))
  d_content <- rawToChar(result$content)
  print(d_content)
  d_content <- fromJSON(d_content)
  return(d_content)
}
```

## Creating data url and credentials variables

```{r, ehco = F, message = F, warning= F}
kobo_server_url<-"https://kobo.humanitarianresponse.info/"
url <-"https://kc.humanitarianresponse.info/api/v1/data.csv"
kobo_user = "rcop_test"
kobo_pw = "e4#xV8zKAU)A)h58"
list_forms <- as.data.frame(kobohr_getforms_csv (url,kobo_user, kobo_pw))

####  Create a variable with the form id and create to dataurl link
form_id = "965004"
asset_uid = "a9XVuozs7VQ2bvbvMwKSdS"
dataurl<- paste0("https://kc.humanitarianresponse.info/api/v1/data/",form_id,".csv")
```

## Fetching the data from the server

```{r, eval = F}
#### Download data as csv
df <-  download_data(dataurl,kobo_user,kobo_pw)

#### Create an export and download it
type <- "xls"
lang <- "xml"
fields_from_all_versions <- "TRUE"
hierarchy_in_labels <- "FALSE"
group_sep = "/"

d_exports<-create_export(type=type,
                                lang=lang,
                                fields_from_all_versions=fields_from_all_versions,
                                hierarchy_in_labels=hierarchy_in_labels,
                                group_sep=group_sep,
                                asset_uid=asset_uid,
                                kobo_user,
                                kobo_pw)

result<-httr::GET (url=paste0(as.character(d_exports$url),"?format=json"),
                    authenticate(kobo_user,kobo_pw),
                    progress()
)


result_file<-httr::GET (url=jsonlite::fromJSON(rawToChar(result$content))$result,
                   authenticate(kobo_user,kobo_pw),
                   progress()
)

httr::GET (result_file$url,
           authenticate(kobo_user,kobo_pw),
           progress(),
           write_disk(tf <- tempfile(fileext = ".xlsx"))
)

df2 <- read_excel(tf)
```

