# Cleaning

## Time Check - based on start end time
Required Libraries: "dplyr" and "lubridate"

****

"time_check" will return the elapsed time for each interview based on it's start and end columns also classifies if it's "too short", "too long", or "okay".
As an example, we will use a dummy data set to apply the function to it. The function also needs a time_min (the minimum time in minutes that an interview should take to be completed) and a time_max (the maximum time in minutes that an interview should take to be completed) parameters.
```{r, eval=FALSE}
library(dplyr)
library(lubridate)

# Creating a dummy data set
start <- c("2020-12-01T09:40:05.750+04:30","2020-12-01T09:40:18.709+04:30","2020-12-01T09:45:40.879+04:30","2020-12-01T09:46:28.328+05:00")
end <- c("2020-12-01T09:44:44.438+04:30","2020-12-01T10:01:27.890+04:30","2020-12-01T10:17:44.021+04:30","2020-12-01T10:18:32.717+05:00")

my_df <- data.frame(start, end)

# Initializing variables
time_min <- 8
time_max <- 30

# declaring the function
time_check <- function(df, time_min, time_max){
  df <- df %>% mutate(interview_duration = difftime(as.POSIXct(ymd_hms(end)), as.POSIXct(ymd_hms(start)), units = "mins"),
                      CHECK_interview_duration = case_when(
                        interview_duration < time_min ~ "Too short",
                        interview_duration > time_max ~ "Too long",
                        TRUE ~ "Okay"
                      )
  )
  return(df)
}

# Applying the function to data frame
processed_df <- time_check(my_df, time_min, time_max)
  
```


****

## Time Check - based on audit files and start end time
Required Libraries: "dplyr", "lubridate", and "expss"

****

"time_check_audit" will calculate the interview duration using the audit files. And, if the audit file for that particular uuid is not found in the audit directory (where you paste the audit files), it will calculate it using start and end time columns in the data set.
Audit files should be stored inside a folder (to avoid redefining its name while calling the function, call it "audit_files") in the project folder.
```{r, eval=FALSE}
library(dplyr)
library(lubridate)
library(expss)

# Creating a dummy data set
start <-  c("2020-12-01T09:40:05.750+04:30","2020-12-01T09:40:18.709+04:30","2020-12-01T09:45:40.879+04:30","2020-12-01T09:46:28.328+05:00")
end <-  c("2020-12-01T09:44:44.438+04:30","2020-12-01T10:01:27.890+04:30","2020-12-01T10:17:44.021+04:30","2020-12-01T10:18:32.717+05:00")
uuid <-  c("uuid_1","uuid_2","uuid_3","uuid_4")

my_df <- data.frame(start, end, uuid)

# Initializing variables
time_min <- 8
time_max <- 30


# declaring the function
time_check_audit <- function(df_raw, x_uuid="_uuid", time_min, time_max, audit_dir_path = "./audit_files/", today = "today"){
  if (!any(duplicated(df_raw[[x_uuid]]))) {
    # Audit Checks
    audit_dir_path < -audit_dir_path
    uuids <- dir(audit_dir_path)
    dfl <- list()
    all_uuids <- length(uuids)
    
    for(i in 1: length(uuids)){
      df <- read.csv(paste0(audit_dir_path, uuids[i], "/audit.csv"))
      df <- df %>% filter(node != "")
      duration_ms <- sum(df$end - df$start)
      duration_secs <-duration_ms/1000
      duration_minutes <- round(duration_secs/60,1)
      dfl[[i]] <- data.frame(uuid = uuids[i], duration_ms=duration_ms, 
                             durations_secs=duration_secs, duration_minutes = duration_minutes)
      cat("\014","Running audit: ", round((i/all_uuids) * 100,0),"%\n", sep = "")
    }
    duration_df <- do.call("rbind", dfl)
    duration_df <- dplyr::rename(duration_df, `_uuid` = uuid)
    
    
    #time check based on start end time
    df_str_audit_all <- df_raw %>% 
      mutate(start_end = difftime(as.POSIXct(ymd_hms(end)), as.POSIXct(ymd_hms(start)), units = "mins"))

    # df_str_audit_all <- df_raw %>%
    #   left_join(select(df_no_audit_files, x_uuid, start_end), by = c("_uuid"="_uuid"))
    
    # Join Audit checks and main data set
    df_str_audit_all <- df_str_audit_all %>%
      left_join(select(duration_df, x_uuid, duration_minutes), by = c("_uuid"="_uuid"))

    # Checking time duration with audit file, if not available, from dataset start/end.
    df_str_audit_all <- df_str_audit_all %>%
      mutate(interview_duration = if_na(duration_minutes, start_end),
             CHECK_interview_duration = case_when(
               interview_duration < time_min ~ "Too short",
               interview_duration > time_max ~ "Too long",
               TRUE ~ "Okay")
      ) %>% select( -c(duration_minutes,start_end))
    
    return(df_str_audit_all)
  }else{
    stop("Error: df_raw has duplicate in uuid column, resolve the duplication to proceed!")
  }
}

# Applying the function to data frame
processed_df <- time_check_audit(my_df, time_min, time_max)
```


## Time Check - calculating the elapsed time between each interview
Required Libraries: lubridate

****

"time_btwn_ints" will calculate the elapsed time between the ending time of the first interview of an enumerator and the start time of its second interview, and the process applies for all interviews of each enumerator.
It needs a location identifier as a parameter to check if the elapsed time is matching with the threshold (given as a parameter) in the same location or not. Also it will be checked if the elapsed time matches the given threshold for interviews in different locations by the same enumerator.
```{r, eval=FALSE}
library(lubridate)

# Creating a dummy data set
start <-  c("2020-12-01T09:40:05.750+04:30",
            "2020-12-01T09:45:18.709+04:30",
            "2020-12-01T09:45:40.879+04:30",
            "2020-12-01T10:02:40.879+04:30",
            "2020-12-01T10:25:28.32+04:30")
end <-  c("2020-12-01T09:44:44.438+04:30",
          "2020-12-01T10:01:27.890+04:30",
          "2020-12-01T10:17:44.021+04:30",
          "2020-12-01T10:05:40.879+04:30",
          "2020-12-01T10:38:32.717+04:30")
device_id <- c("000215",
               "000215",
               "000216",
               "000215",
               "000216")
village <- c("Village A",
             "Village A",
             "Village B",
             "Village A",
             "Village C")

my_df <- data.frame(start, end, device_id, village)

# declaring the function
time_btwn_ints <- function(df, device_id, start_col = "start", end_col = "end", 
                           village_col, same_village_threshold = 3, diff_village_threshold = 5){
  checked_df <- df

  # sort by device_id and start_col
  checked_df <- checked_df[order(checked_df[[start_col]]), ]
  checked_df <- checked_df[order(checked_df[[device_id]]), ]

  # For each row starting from the second row:
  # 1) calculate the time between the end of the (r-1) survey and the start of the (r) survey
  # 2) insert the eight check-message based on the calculated time and the village
  issue.same.village <- paste0("The elapsed time between two interviews in the same village is less than ",same_village_threshold, " minutes")
  issue.diff.village <- paste0("The elapsed time between two interviews in different villages is less than ", diff_village_threshold, " minutes")
  checked_df$check <- "OK"
  checked_df$gap_between_ints <- NA
  for (r in 2:nrow(checked_df)){
    if (as.character(checked_df[r, device_id])==as.character(checked_df[r-1, device_id])){
      checked_df$gap_between_ints[r] <- difftime(as.POSIXct(ymd_hms(checked_df[r, start_col])),
                                                 as.POSIXct(ymd_hms(checked_df[r-1, end_col])),
                                                 units = "mins")

      if (as.character(checked_df[r, village_col])==as.character(checked_df[r-1, village_col])){
        if (checked_df$gap_between_ints[r] < same_village_threshold) checked_df[r, "check"] <- issue.same.village
      } else{
        if (checked_df$gap_between_ints[r] < diff_village_threshold) checked_df[r, "check"] <- issue.diff.village
      }
    }
  }

  return(checked_df)
}

# Applying the function to data frame
processed_df <- time_btwn_ints(df = my_df, device_id = "device_id",village_col = "village", same_village_threshold = 2,diff_village_threshold = 10)

```

****

## Data cleaning - Clean data based on cleaning log
Required Libraries: Base R Packages

****

The function "incorporate_logs" applies cleaning log on raw data. It gets cleaning log in a format that must contain 5 mandatory columns including uuid, question.name, old.value, new.value, and changed. as result it would return the cleaned data frame, master cleaning log (logs that are both applied and not applied on data), the version of cleaning log that was applied on raw data, a report of duplicate logs, and logs that their question name or uuid is not available in raw data frame.

****
```{r, eval=FALSE}
# Creating a dummy data set
city_name <- c("kabul", "new dlehi", "peshawar","new york")
population <- c(4430000, 21750000, 1970000, 8419000)
uuid <- c("eae001", "eae002", "eae003","eae004")

my_df <- data.frame(city_name, population, uuid)

# Creating a dummy cleaning log
old.value <- c("kabul", 4430000, "europe","kabul")
question.name <- c("city_name", "population", "continent","city_name")
new.value <- c("moscow",11920000,"asia","moscow")
uuid <- c("eae001","eae001","eae001","eae001")
changed <- c("yes","yes","yes","yes")

cleaning_log <- data.frame(uuid,question.name, old.value, new.value, changed)

### declaring the function
incorporate_logs = function(raw_df, cleaning_log, df_group_seprator = "/", uuid_col = "_uuid"){
  error <- "Error!
Execution was haulted due to one of the following issues:
  - Cleaning log is empty
  - There is no changes in data (in cleaning log changed property for all logs is set to 'NO')
  - One/morethan one of the (uuid, question.name, old.value, new.value, and changed) columns are missing or column names are misspelled
"
  if (sum(grepl("uuid|question.name|old.value|new.value|changed", names(cleaning_log)))==5) {
    `%nin%` = Negate(`%in%`)
    # changing the group seprator (operator) from "/" to "."
    names(raw_df) <- gsub(df_group_seprator,".",names(raw_df))
    cleaning_log$question.name <- gsub(df_group_seprator,".", cleaning_log$question.name)
    
    # subsetting logs that their question is not (available) in dataset
    logs_not_in_rawdf <- cleaning_log[cleaning_log$question.name %nin% names(raw_df) | cleaning_log$uuid %nin% raw_df[[uuid_col]], ]
    logs_not_in_rawdf <- logs_not_in_rawdf[logs_not_in_rawdf$changed %in% c("yes","Yes"),]
    
    # subsetting logs that their question exist in raw data frame and its new value is changed
    cleaning_log.changed <- cleaning_log[cleaning_log$question.name %in% names(raw_df) & cleaning_log$uuid %in% raw_df[[uuid_col]], ]
    cleaning_log.changed <- cleaning_log.changed[cleaning_log.changed$changed %in% c("yes","Yes"),]
    
    # capturing duplicate logs
    cleaning_log$unique_key <- paste(cleaning_log$uuid, cleaning_log$question.name, sep = "_")
    duplicate_logs <- cleaning_log[(duplicated(cleaning_log$unique_key) | duplicated(cleaning_log$unique_key, fromLast = T)),]
    
    # cleaning master cleaning log
    cleaning_log <- cleaning_log[cleaning_log$uuid %nin% logs_not_in_rawdf$uuid | cleaning_log$question.name %nin% logs_not_in_rawdf$question.name,]
    cleaning_log <- cleaning_log[!is.na(cleaning_log$question.name), ]
    cleaning_log <- cleaning_log[!is.na(cleaning_log$uuid), ]
    
    raw_df_valid <- raw_df
    if (nrow(cleaning_log.changed)>0) {
      # Apply cleaning log on raw data
      for (rowi in 1:nrow(cleaning_log.changed)){
        uuid_i <- cleaning_log.changed$uuid[rowi]
        var_i <- cleaning_log.changed$question.name[rowi]
        old_i <- cleaning_log.changed$old.value[rowi]
        new_i <- cleaning_log.changed$new.value[rowi]
        if(class(raw_df_valid[[var_i]]) == "character"){
          new_i <- as.character(new_i)
        }else if(class(raw_df_valid[[var_i]]) == "numeric"){
          new_i <- as.numeric(new_i)
        }else if(class(raw_df_valid[[var_i]]) == "logical"){
          new_i <- as.integer(new_i)
        }else if(class(raw_df_valid[[var_i]]) == "integer"){
          new_i <- as.integer(new_i)
        }
        # Find the variable according to the row of the cleaning log
        raw_df_valid[raw_df_valid[[uuid_col]] == uuid_i, var_i] <- new_i
        print(paste(rowi,"uuid:", uuid_i, "Old value:", old_i, "changed to", new_i, "for", var_i))
      }
      return(list(cleaned_df = raw_df_valid, cleaning_log.applied = cleaning_log.changed, logs_not_in_rawDF = logs_not_in_rawdf, duplicate_logs = duplicate_logs, master_cleaning_log = cleaning_log))
    }else{
      cat(error)
      return(list(cleaned_df = raw_df_valid, cleaning_log.applied = cleaning_log.changed,logs_not_in_rawdf = logs_not_in_rawdf))
    }
  }else{
    cat(error)
  }
}

### Applying the function to data frame
incorprated_logs <- incorporate_logs(my_df, cleaning_log, uuid_col = "uuid")

cleaned_data <- incorprated_logs$cleaned_df
master_cleaning_log <- incorprated_logs$master_cleaning_log
logs_not_in_rawDf <- incorprated_logs$logs_not_in_rawDF
cleaning_log.applied <- incorprated_logs$cleaning_log.applied
duplicate_log <- incorprated_logs$duplicate_logs

```

****

kobold/ butteR/ auditCheckR

## Example one

## Example two

