# Composite indicators
composeR
dplyr::case_when

## Composite with 1 categorical variable - select one

This example will look at creating an improved source of water variable. The indicator of interest is **f11_dinking_water_source** and the options for un-improved source of water are:

- trucked_in_water_truck_with_a_tank_etc
- drinking_water_from_water_kiosk_booth_with_water_for_bottling
- bottled_water_water_purchased_in_bottles
- other_specify

```{r }
#Creating a vector of un-improved source of water
unimproved_source <- c("trucked_in_water_truck_with_a_tank_etc", "drinking_water_from_water_kiosk_booth_with_water_for_bottling", "bottled_water_water_purchased_in_bottles", "other_specify")

#Using base R
main_dataset$wash_drinkingwater_improved_source_baser <- ifelse(main_dataset$f11_dinking_water_source %in% unimproved_source, "not_improved", "improved")
table(main_dataset$f11_dinking_water_source, main_dataset$wash_drinkingwater_improved_source_baser, useNA = "ifany")

#Using base R 
main_dataset$wash_drinkingwater_improved_source_baser2 <- NA
main_dataset$wash_drinkingwater_improved_source_baser2[main_dataset$f11_dinking_water_source %in% unimproved_source] <- "not_improved"
main_dataset$wash_drinkingwater_improved_source_baser2[!(main_dataset$f11_dinking_water_source %in% unimproved_source)] <- "improved"

#Same same
table(main_dataset$wash_drinkingwater_improved_source_baser, main_dataset$wash_drinkingwater_improved_source_baser2, useNA = "ifany")

#Using case_when
main_dataset <- main_dataset %>%
  mutate(wash_improved_source_dplyr = case_when(f11_dinking_water_source %in% unimproved_source ~ "not_improved",
                                                TRUE ~ "improved"))
table(main_dataset$f11_dinking_water_source, main_dataset$wash_improved_source_dplyr, useNA = "ifany")

#Same same
table(main_dataset$wash_drinkingwater_improved_source_baser, main_dataset$wash_improved_source_dplyr)
```

In the previous example, there was no missing value. For this example, the indicator to be built will turn a yes/no question into a dummy variable (1 and 0). The variable of interest is  *b16_hohh_pension_eligible* 

```{r }
table(main_dataset$b16_hohh_pension_eligible, useNA = "ifany")

#Using base R
main_dataset$hohh_pension_eligible_dummy_baser <- ifelse(main_dataset$b16_hohh_pension_eligible == "yes", 1, 0)

table(main_dataset$b16_hohh_pension_eligible, main_dataset$hohh_pension_eligible_dummy_baser, useNA = "ifany")

#Using case_when
main_dataset <- main_dataset %>%
  mutate(hohh_pension_eligible_dummy_dplyr = case_when(b16_hohh_pension_eligible == "yes" ~ 1,
                                                b16_hohh_pension_eligible == "no" ~ 0))

table(main_dataset$b16_hohh_pension_eligible, main_dataset$hohh_pension_eligible_dummy_dplyr, useNA = "ifany")

#Same same
table(main_dataset$hohh_pension_eligible_dummy_baser, main_dataset$hohh_pension_eligible_dummy_dplyr, useNA = "ifany")
```

```{r }
#Watch out for NA. This was is not correct. 
main_dataset <- main_dataset %>%
  mutate(hohh_pension_eligible_dummy_dplyr2 = case_when(b16_hohh_pension_eligible == "yes" ~ 1,
                                                TRUE ~ 0))
table(main_dataset$b16_hohh_pension_eligible, main_dataset$hohh_pension_eligible_dummy_dplyr2, useNA = "ifany")

#Not same same
table(main_dataset$hohh_pension_eligible_dummy_baser, main_dataset$hohh_pension_eligible_dummy_dplyr2, useNA = "ifany")
```


## Composite with 2 categorical variables 
This example will look at creating an indicator whether or not the sources for drinking and for cooking, cleaning and non-drinking purposes are both improved. The indicators of interest are **f11_dinking_water_source** (and more specifically **wash_drinkingwater_improved_source_baser** from previous paragraph) and **f14_technical_water_source** (*F14_What is your HH's main source of water for cooking, cleaning, and non-drinking purposes*). 

First, a new variable has to be created, **wash_otherwater_improved_source_baser**.

```{r }
#Using base R
main_dataset$wash_otherwater_improved_source_baser <- ifelse(main_dataset$f14_technical_water_source %in% unimproved_source, "not_improved", "improved")

main_dataset$wash_bothwater_improved_source_baser <- ifelse(main_dataset$wash_drinkingwater_improved_source_baser == "improved" & main_dataset$wash_otherwater_improved_source_baser == "improved", "both_improved", "not_both_improved")

table(main_dataset$wash_drinkingwater_improved_source_baser, main_dataset$wash_otherwater_improved_source_baser, main_dataset$wash_bothwater_improved_source_baser, useNA = "ifany")
```

Now, the variable will be coded to have 3 categories instead: both improved, at least drinking water and not improved.

```{r }
main_dataset$wash_bothwater_improved_source_baser2 <- ifelse(main_dataset$wash_drinkingwater_improved_source_baser == "improved" & main_dataset$wash_otherwater_improved_source_baser == "improved", "both_improved", 
                                                             ifelse(main_dataset$wash_drinkingwater_improved_source_baser == "improved", "at_least_drinking", "not_both_improved"))

table(main_dataset$wash_drinkingwater_improved_source_baser, main_dataset$wash_otherwater_improved_source_baser, main_dataset$wash_bothwater_improved_source_baser2, useNA = "ifany")
```

```{r }
#Using dplyr
main_dataset <- main_dataset %>%
  mutate(wash_bothwater_improved_source_dplyr = case_when(wash_drinkingwater_improved_source_baser == "improved" & wash_otherwater_improved_source_baser == "improved" ~ "both_improved",
                                                          wash_drinkingwater_improved_source_baser == "improved" ~ "at_least_drinking", 
                                                          TRUE ~ "not_both_improved"
                                                          ))

table(main_dataset$wash_bothwater_improved_source_dplyr,  main_dataset$wash_bothwater_improved_source_baser2, useNA = "ifany")
```
## Composite with 1 categorical variable - select multiple
In this example we are creating an indicator to score whether or not drinking water is being processed or purified before usage. The indicator is calculated based on one categorical question **f12_drinking_water_treat**

```{r }
#Using dplyr
main_dataset <- main_dataset %>%
   mutate(
     wash_indicator1 = case_when(
       f12_drinking_water_treat.do_not_process_purify == 1 ~ 3 ,
       f12_drinking_water_treat.cleaning_with_chemicals_chlorination == 1 |
       f12_drinking_water_treat.water_precipitation == 1 |
       f12_drinking_water_treat.filtering_the_water_pitcher_filter == 1 |
       f12_drinking_water_treat.filtering_the_water_reverse_osmosis_filter == 1 |
       f12_drinking_water_treat.boiling == 1 |
       f12_drinking_water_treat.percolation == 1 ~ 1
   ) )


#Using base R
main_dataset$wash_indicator2 = ifelse(
  main_dataset$f12_drinking_water_treat.do_not_process_purify == 1,3,
  ifelse(
    main_dataset$f12_drinking_water_treat.cleaning_with_chemicals_chlorination == 1 |
       main_dataset$f12_drinking_water_treat.water_precipitation == 1 |
       main_dataset$f12_drinking_water_treat.filtering_the_water_pitcher_filter == 1 |
       main_dataset$f12_drinking_water_treat.filtering_the_water_reverse_osmosis_filter == 1 |
       main_dataset$f12_drinking_water_treat.boiling == 1 |
       main_dataset$f12_drinking_water_treat.percolation == 1,
    1,NA))
```

## Composite with 1 numerical variable
This example will look at creating one a categorical variable based on a number, **f6_how_many_wood_hh_consumed_last_winter  **, 'less than 5', between 5 (included) and 10 and '10 and above'.

```{r }
main_dataset <- main_dataset %>%
  mutate(wood_consumed_categories = case_when(f6_how_many_wood_hh_consumed_last_winter < 5 ~ "less_than_5",
                                              f6_how_many_wood_hh_consumed_last_winter < 10 ~ "between5_and10",
                                              f6_how_many_wood_hh_consumed_last_winter >= 10 ~ "ten_above"))

```

## Composite with 2 numerical variables

This example will look at creating one of the indicators necessary to compute the FCS. In some cases, we need to check if the sum of number of days for 2 types of food are above 7 or not. If the sum is above 7, then it has to return 7 otherwise, the sum of both variables.

```{r }
# Combine cereals/roots and meat/eggs and make maximum 7 days
## Using base R
main_dataset$fcs_cereal_roots  <- ifelse((main_dataset$g1_cereals_consumption + main_dataset$g2_roots_consumption) > 7,
                                7,
                                main_dataset$g1_cereals_consumption + main_dataset$g2_roots_consumption)

## Using dplyr
main_dataset <- main_dataset %>%
  mutate(fcs_meat_eggs = ifelse((g5_meat_consumption + g6_eggs_consumption) > 7, 7,
                                 g5_meat_consumption + g6_eggs_consumption))


```

## Composite with 2 or more numerical variables


This example will look at creating the food consumption score. 
```{r }
main_dataset <- main_dataset %>%
  mutate(FCS_score_dplyr = fcs_cereal_roots * 2 + g3_vegetables_consumption * 1 + g4_fruits_consumption * 1 + fcs_meat_eggs * 4 + g7_pulses_consumption * 3 + g8_dairy_consumption * 4 + g9_oil_consumption * 0.5 + g10_sugar_consumption * 0.5)

main_dataset$FCS_score_baser  <- (main_dataset$fcs_cereal_roots * 2)+
  (main_dataset$g3_vegetables_consumption * 1)+
  (main_dataset$g4_fruits_consumption * 1)+
  (main_dataset$fcs_meat_eggs * 4)+
  (main_dataset$g7_pulses_consumption * 3)+
  (main_dataset$g8_dairy_consumption * 4)+
  (main_dataset$g9_oil_consumption * 0.5)+
  (main_dataset$g10_sugar_consumption * 0.5)

```


## Household Hunger Score (HHS) 

For this example we will use addindicators package to compute the HHS 
Link to the package: https://github.com/impact-initiatives/addindicators
The package can be installed using devtools::install_github("impact-initiatives/addindicators")

```{r}

# Align the coding of the frequencies
main_dataset <- main_dataset %>% mutate(
  h3_1_how_often_day_and_night_hungry = case_when(
    h3_1_how_often_day_and_night_hungry == "1_rarely_12" ~ "1_rarely_1_2",
    TRUE ~ h3_1_how_often_day_and_night_hungry
  )
)


main_dataset <- addindicators::add_hhs(
  main_dataset,
  hhs_nofoodhh_1 = "h1_nofood_noresources",
  hhs_nofoodhh_1a = "h1_1_how_often_nofood_noresources",
  hhs_sleephungry_2 = "h2_sleep_hungry",
  hhs_sleephungry_2a = "h2_1_how_often_sleep_hungry",
  hhs_alldaynight_3 = "h3_day_and_night_hungry",
  hhs_alldaynight_3a = "h3_1_how_often_day_and_night_hungry",
  yes_answer = "1_yes",
  no_answer = "0_no",
  rarely_answer = "1_rarely_1_2",
  sometimes_answer = "2_sometimes_3_10",
  often_answer = "3_often_10plus_times"
)

main_dataset %>% select(hhs_cat) %>% head()

```

## Food Cconsumption Score (FCS) 

For this example we will use addindicators package to compute the FCS 
Link to the package: https://github.com/impact-initiatives/addindicators
The package can be installed using devtools::install_github("impact-initiatives/addindicators")


```{r}

main_dataset <- addindicators::add_fcs(
  main_dataset,
  cutoffs = "normal 21.5-35", # "normal 21.5-35", or "alternative 28-42"
  fcs_cereal = "g1_cereals_consumption",
  fcs_legumes = "g2_roots_consumption" ,
  fcs_veg = "g3_vegetables_consumption",
  fcs_fruit = "g4_fruits_consumption",
  fcs_meat = "g5_meat_consumption",
  fcs_dairy = "g8_dairy_consumption",
  fcs_sugar = "g10_sugar_consumption",
  fcs_oil = "g9_oil_consumption"
)

main_dataset %>% select(fcs_score,fcs_cat) %>% head()

```


## Livelihood coping strategy index (LCSI) 

For this example we will use addindicators package to compute the LCSI 
Link to the package: https://github.com/impact-initiatives/addindicators
The package can be installed using devtools::install_github("impact-initiatives/addindicators")

```{r}


main_dataset <- addindicators::add_lcsi(main_dataset,
                     lcsi_stress_vars = c("i1_coping_strategies", "i2_spent_savings", "i3_borrowed_food", "i4_live_another_family") ,
                     lcsi_crisis_vars = c("i6_removed_children_from_school_kindergarten", "i7_reduced_health_care", "i8_reduced_education") ,
                     lcsi_emergency_vars = c("i9_sold_house", "i10_hh_migrated", "i11_moved_for_work"),
                     yes_val = "yes",
                     no_val = "no_but_needed",
                     exhausted_val = "no_already_done_so_cant_do_it_anymore",
                     not_applicable_val = "no_was_not_needed")
  
main_dataset %>% select(lcsi_cat) %>% head()

```


## New indicators from a loop to main dataset
e.g. aggregating the number of children going to school from a loop


## The co-occurrence of needs in an MSNI

The following example will fist create an choice multiple question and then will create a parent column which can be use for calculating co-occurrence of needs.
```{r eval=F }
lsg_cols <- c("wash","education","health","food","protection")

for(i in lsg_cols){
  col_name <- paste0("most_common_profile.",i)
  most_common_data <- most_common_data %>% mutate(
    !!sym(col_name) := as.numeric(!!sym(i)) >2 )}

most_common_data <- most_common_data %>% mutate(
  most_common_profile = NA)

most_common_data <- most_common_data %>% 
  illuminate::recalculate_concerted_col_for_select_multiple(uuid = "uuid")

```


