# Analysis



## srvyr package

"The srvyr package aims to add dplyr like syntax to the survey package." It is a very useful package for a variety of aggregations of survey data.

```{r, tidy=FALSE, message= F, warning=F, error=F, echo=T}
###makes some additions. 
library(tidyverse)
library(srvyr)
library(kableExtra)
df<-read_csv("inputs/UKR2007_MSNA20_HH_dataset_main_rcop.csv")
dfsvy<-as_survey(df)
```

### Categorical variables
srvyr package allows categorical variables to be broken down using a similar syntax as dplyr.  Using dplyr you might typically calculate a percent mean as follows:
```{r}

df %>% 
  group_by(b9_hohh_marital_status) %>% 
  summarise(
    n=n()
  ) %>% 
  ungroup() %>% 
  mutate(
    pct_mean=n/sum(n)
  )

```

To calculate the percent mean of a categorical variable using srvyr object is required. The syntax is quite similar to dplyr, but a bit less verbose. By specifying the vartype as "ci" we also get the upper and lower confidence intervals

```{r}
dfsvy %>% 
  group_by(b9_hohh_marital_status) %>% 
  summarise(
    pct_mean = survey_mean(vartype = "ci")
  )

```
To calculate the weigthed percent mean of a multiple response option you need to created a srvyr object including the weights. The syntax is similar to dyplr and allows for the total columns

```{r}
weighted_object <- main_dataset %>% as_survey_design(ids = X_uuid, weights = stratum.weight)

weighted_table <- weighted_object %>% 
                  group_by(adm1NameLat) %>% #group by categorical variable
                  summarise_at(vars(starts_with("b10_hohh_vulnerability.")), survey_mean) %>% #select multiple response question
                  ungroup() %>% 
                  bind_rows(
                            summarise_at(weighted_object,
                            vars(starts_with("b10_hohh_vulnerability.")), survey_mean) # bind the total
                  ) %>%
                mutate_at(vars(adm1NameLat), ~replace(., is.na(.), "Total")) %>% 
                select(., -(ends_with("_se"))) %>%  #remove the colums with the variance type calculations
                mutate_if(is.numeric, ~ . * 100) %>%
                mutate_if(is.numeric, round, 2)


print(weighted_table)


```
### Numeric variables
srvyr treats the calculation/aggregation of numeric variables differently in an attempt to mirror dplyr syntax 

to calculate the mean and median expenditure in dplyr you would likely do the following
```{r}

df %>% 
  summarise(
    mean_expenditure= mean(n1_HH_total_expenditure,na.rm=T),
    median_expenditure=median(n1_HH_total_expenditure,na.rm=T),
    )

```

If you wanted to subset this by another variable in dplyr you would add the group_by argument
```{r}

df %>% 
  group_by(strata) %>% 
  summarise(
    mean_expenditure= mean(n1_HH_total_expenditure,na.rm=T),
    median_expenditure=median(n1_HH_total_expenditure,na.rm=T),
    )

```

This is the reason why the syntax also varies between categorical and numeric variables in srvyr. Therefore, to do the same using srvyr you would do the following (with a survey object). Note that due to this difference in syntax the na.rm argument works for numeric variables, but **does not work** for categorical variables. This was modified when srvyr was updated from v 0.3.8

```{r}
dfsvy %>% 
  summarise(
   mean= survey_mean(n1_HH_total_expenditure,na.rm=T,vartype = "ci"),
  )


```

similar to dplyr you can easily add a group_by argument to add a subset calculation
```{r}
dfsvy %>% 
  group_by(strata) %>% 
  summarise(
   mean= survey_mean(n1_HH_total_expenditure,na.rm=T,vartype = "ci"),
  )


```



### select_one 
### select_mutiple

## Analysis with numerical variables
### Averages
###Summarytools (CRAN package)

###hypegrammaR / koboquest / butteR

### Median
####Spatstat
[Spatstat](https://cran.r-project.org/web/packages/spatstat/index.html) - library with set of different functions for analyzing Spatial Point Patterns but also quite useful for analysis of weighted data.

At first let's select all numerical variables from the dataset using Kobo questionnaire and dataset. It can be done with the following custom function:
```{r}
library(spatstat)

select_numerical <- function(dataset, kobo){
  kobo_questions <- kobo[grepl("integer|decimal|calculate", kobo$type),c("type","name")]
  names.use <- names(dataset)[(names(dataset) %in% as.character(kobo_questions$name))]
  numerical <- dataset[,c(names.use,"X_uuid",'strata','stratum.weight')] #Here we can select any other relevant variables
  numerical[names.use] <- lapply(numerical[names.use], gsub, pattern = 'NA', replacement = NA)
  numerical[names.use] <- lapply(numerical[names.use], as.numeric)
  return(numerical)
}

numerical_questions <- select_numerical(main_dataset, questions)

numerical_classes <- sapply(numerical_questions[,1:c(ncol(numerical_questions)-3)], class) #Here we can check class of each selected variable
numerical_classes <- numerical_classes["character" %in% numerical_classes] #and here we check if any variable has class "character"
numerical_questions <- numerical_questions[ , !names(numerical_questions) %in% numerical_classes] #if any variable has a character class then we remove it
rm(numerical_classes)#and here we removing vector with classes from our environment

```

Now let's calculate weighted median for "n1_HH_total_expenditure".
```{r}
weighted.median(numerical_questions$n1_HH_total_expenditure, numerical_questions$stratum.weight, na.rm=TRUE)
```

But if we want to calculate weighted medians for each variable we will need to iterate this function on those variables. But first we will need to exclude variables with less than 3 observations.
```{r}
counts <- numerical_questions %>%
select(-X_uuid, -strata) %>%
summarise(across(.cols = everything(), .fns= ~sum(!is.na(.)) ))%>%
t()#Calculating count of observation for each variable

numerical_questions <- numerical_questions[ , (names(numerical_questions) %in% rownames(subset(counts, counts[,1] > 3)))]
#removing variables with less than 3 observations


medians <- lapply(numerical_questions[1:46], weighted.median, w = numerical_questions$stratum.weight, na.rm=TRUE)%>%
  as.data.frame()
```

Now we can transpond resulting vector and add description to the calculation
```{r}
medians <- as.data.frame(t(medians),stringsAsFactors = FALSE)
names(medians)[1] <- "Median_wght"
head(medians)
```

## Ratios

This function computes the % of answers for select multiple questions.\
The output can be used as a data merge file for indesign.\
The arguments are: dataset, disaggregation variable, question to analyze, and weight column.\

### Example1: \
**Question**: e11_items_do_not_have_per_hh\
**Disaggregation variable**: b9_hohh_marital_status\
**weights column**: stratum.weight\

```{r}

ratios_select_multiple <-
  function(df, x_var, y_prefix, weights_var) {
    df %>%
      group_by({
        {
          x_var
        }
      }) %>% filter(!is.na(y_prefix)) %>%
      summarize(across(starts_with(paste0(y_prefix, ".")),
                       list(pct = ~ weighted.mean(., {
                         {
                           weights_var
                         }
                       }, na.rm = T))))
  }

res <-
  ratios_select_multiple(
    main_dataset,
    b9_hohh_marital_status,
    "e11_items_do_not_have_per_hh",
    stratum.weight
  )

head(res)
```

### Example2 - Analyzing multiple questions and combine the output:\
**Question**: e11_items_do_not_have_per_hh and b11_hohh_chronic_illness\
**Disaggregation variable**: b9_hohh_marital_status\
**weights column**: stratum.weight\

```{r}
res <-
  do.call("full_join", map(
    c("e11_items_do_not_have_per_hh", "b11_hohh_chronic_illness"),
    ~ ratios_select_multiple(main_dataset, b9_hohh_marital_status, .x, stratum.weight)
  ))

head(res)

```

### Example3 - No dissagregation is needed and/or the data is not weighted:\
**Question**: e11_items_do_not_have_per_hh and b11_hohh_chronic_illness\
**Disaggregation variable**: NA\
**weights column**: NA\

```{r}


main_dataset$all <- "all"
main_dataset$no_weights <- 1

res <-
  do.call("full_join", map(
    c("e11_items_do_not_have_per_hh", "b11_hohh_chronic_illness"),
    ~ ratios_select_multiple(main_dataset, all, .x, no_weights)
  ))

head(res)

```



## Weights

Weights can be used for different reasons. In most of cases, we will use weights to correct for difference between strata size. Once the population has a certain (big) size, for the same design, the size of the sample won't change much. In the case of the dataset, there are 4 stratas:\
- within 20km of the border in rural areas : 89,408 households\
- within 20km of the border in urban areas : 203,712 households\
- within 20km of the border in rural areas : 39,003 households\
- within 20km of the border in rural areas : 211,857 households

Using any sampling calculator, for a 95% confindence level, 5% margin of error and 5% buffer, we will have around 400 samples per strata even though the urban areas are much bigger.\
*Note: The error of 5% for 90,000 is 4,500 households while for 200,000 households is 10,000 households*

If we look at each strata and the highest level of education completed by the household head (**b20_hohh_level_education**), we can see that the percentage of household were the head of household completed higher education varies between between 7 to 13 % in each strata.

```{r, echo = F}
main_dataset %>% 
  group_by(strata, b20_hohh_level_education) %>% 
  tally() %>% 
  mutate(tot_n = sum(n),
         prop = round(n/tot_n*100)) %>% 
  filter(b20_hohh_level_education == "complete_higher")
```

However, if we want to know overall the percentage of who finished higher education we cannot just take the overall percentages, i.e. $\frac{40 + 51 +38 + 28}{407 + 404 + 402 + 404}$ = $\frac{157}{1617}$ = 10%.\
We cannot do this because the first strata represent 90,000 households, the second 200,000 households, the third 40,000 households and the last one 210,000 households. We will use weights to compensate this difference.

We will use this formula:\
weight of strata = $\frac{\frac{strata\\ population}{total\\ population}}{\frac{strata\\ sample}{total\\ sample}}$

### tidyverse

The following example will show how to calculate the weights with **tidyverse**.

To calculate weights, we will need all the following information: - *population of the strata*,\
- *total population*,\
- *sample* and\
- *total sample*.\
The population information should come from the sampling frame that was used to draw the sampling.

```{r, echo = T}
my_sampling_frame <- readxl::read_excel("inputs/UKR2007_MSNA20_GCA_Weights_26AUG2020.xlsx", 
                                        sheet = "for_r") %>% 
  rename(population_strata = population)
my_sampling_frame
```

Then, we need to get the actual sample from the dataset.

```{r}
sample_per_strata_table <- main_dataset %>% 
  group_by(strata) %>% 
  tally() %>% 
  rename(strata_sample = n) %>% 
  ungroup()

sample_per_strata_table
```

Then, we can join the tables together and calculate the weights per strata.

```{r, message = F}
weight_table <- sample_per_strata_table %>% 
  left_join(my_sampling_frame) %>% 
  mutate(weights = (population_strata/sum(population_strata))/(strata_sample/sum(strata_sample)))

weight_table
```

Then we can finally add them to the dataset.

```{r}
main_dataset <- main_dataset %>% 
  left_join(select(weight_table, strata, weights), by = "strata")
head(main_dataset$weights)
```

We can check that each strata has only 1 weight.

```{r}
main_dataset %>% 
  group_by(strata, weights) %>%
  tally()
```

We can check that the sum of weights is equal to the number of interviews.

```{r}

sum(main_dataset$weights) == nrow(main_dataset)
```

### surveyweights

**surveyweights** was created to calculate weights. The function **weighting_fun_from_samplingframe** creates a function that calculates weights from the dataset and the sampling frame.

*Note: surveyweights can be found on [github](https://github.com/impact-initiatives/surveyweights).*

First you need to create the weighting function. **weighting_fun_from_samplingframe** will take 5 arguments:\
- sampling.frame: a data frame with your population figures and stratas.\
- data.stratum.column : column name of the strata in the dataset.\
- sampling.frame.population.column : column name of population figures in the sampling frame.\
- sampling.frame.stratum.column : column name of the strata in the sampling frame.\
- data : dataset

```{r}
library(surveyweights)

my_weigthing_function <- surveyweights::weighting_fun_from_samplingframe(sampling.frame = my_sampling_frame,
                                                                      data.stratum.column = "strata",
                                                                      sampling.frame.population.column = "population_strata", 
                                                                      sampling.frame.stratum.column = "strata", 
                                                                      data = main_dataset)
```

See that the **my_weigthing_function** is not a vector of weights. It is a function that takes the dataset as argument and returns a vector of weights.

```{r}
is.function(my_weigthing_function)
```

```{r}
my_weigthing_function
my_weigthing_function(main_dataset) %>% head()
```

*Note: A warning shows that if the column weights exists in the dataset, **my_weigthing_function** will not calculate weights. We need to remove the weights column from the previous example.*

```{r}
main_dataset$weights <- NULL
my_weigthing_function(main_dataset) %>% head()
```

To add the weights, we need to add a new column.

```{r}
main_dataset$my_weights <- my_weigthing_function(main_dataset)

head(main_dataset$my_weights)
```

As for the previous example, we can check that only 1 weight is used per strata.

```{r}
main_dataset %>% 
  group_by(strata, my_weights) %>% 
  tally()
```

We can check that the sum of weights is equal to the number of interviews.

```{r}
sum(main_dataset$my_weights) == nrow(main_dataset)
```


## Repeating the above

## Analysis - top X / ranking (select one and select multiple)

First: select_one question

The indicator of interest is **e1_shelter_type** and the options are:

- solid_finished_house
- solid_finished_apartment
- unfinished_nonenclosed_building
- collective_shelter
- tent
- makeshift_shelter
- none_sleeping_in_open
- other_specify
- dont_know_refuse_to_answer

```{r}
# Top X/ranking for select_one type questions 
select_one_topX <- function(df, question_name, X = 3) { # by default return top 3
  # test message
  if(length(colnames(df)[grepl(question_name, colnames(df), fixed = T)]) == 0) {
    stop(print(paste("question name:", question_name, "doesn't exist in the main dataset. Please double check and make required changes!")))
  }
  
  df <- df %>%
    select(!!sym(question_name))%>%          
    filter(!is.na(!!sym(question_name))) %>% # remove NA values from percentage calculations 
    mutate_at(question_name, as.factor) %>%  # added to have all groups
    group_by(!!sym(question_name), .drop=F) %>% 
    summarise(n = n()) %>%
    mutate(percentages = round(n / sum(n) * 100, digits = 2)) %>%
    arrange(desc(percentages)) %>% 
    mutate_at(question_name, as.character) %>% 
    head(X)   # keep top X percentages only 
  
  if(nrow(df) == 0) warning("None of the choices was selected!")
 
  return(df)
}

# return top 4 shelter types
shelter_type_top4 <- select_one_topX(df = main_dataset, question_name = "e1_shelter_type", X = 4)
```

Second: select_multiple question

The indicator of interest is **e3_shelter_enclosure_issues** and the options are:

- no_issues
- lack_of_insulation_from_cold
- leaks_during_light_rain_snow
- leaks_during_heavy_rain_snow
- limited_ventilation_less_than_05m2_ventilation_in_each_room_including_kitchen
- presence_of_dirt_or_debris_removable
- presence_of_dirt_or_debris_nonremovable
- none_of_the_above
- other_specify
- dont_know

```{r}
# Top X/ranking for select_multiple type questions
select_multiple_topX <- function(df, question_name, X = 3) {
  # test message
  if(length(colnames(df)[grepl(question_name, colnames(df), fixed = T)]) == 0){
    stop(print(paste("question name:", question_name, "doesn't exist in the main dataset. Please double check and make required changes!")))
  }
  
  df <- df %>%
    select(colnames(df)[grepl(paste0(question_name, "."), colnames(df), fixed = T)])
  
  # if question was not answered
  if(all(is.na(df))) warning("None of the choices was selected!")

  # Keeping only the options names in the colnames
  colnames(df) <- gsub(paste0(question_name, "\\."), "", colnames(df))
  
  # calculate top X percentages
  df <- df %>%
    mutate_all(as.numeric) %>% 
    pivot_longer(cols = everything(), names_to = question_name, values_to = "choices") %>%
    group_by(!!sym(question_name), .drop=F) %>% 
    summarise(n = sum(choices, na.rm = T),
              percentages = round(sum(choices, na.rm = T) / n() * 100, digits = 2)) %>% # remove NA values from percentages calculations
    arrange(desc(percentages)) %>% 
    head(X)   # keep top X percentages only 
  
  return(df)
}

# return top 7 shelter enclosure issues
shelter_enclosure_issues_top7 <- select_multiple_topX(df = main_dataset, question_name = "e3_shelter_enclosure_issues", X = 7)
```


## Borda count

## Hypothesis testing
### T-test

#### What is a T-test

A t-test is a type of inferential statistic used to determine if there is a significant difference between the means of two groups. A t-test is used as a hypothesis testing tool, which allows testing of an assumption applicable to a population. 


#### How it works

Mathematically, the t-test takes a sample from each of the two sets and establishes the problem statement by assuming a null hypothesis that the two means are equal. Based on the applicable formulas, certain values are calculated and compared against the standard values, and the assumed null hypothesis is accepted or rejected accordingly.

If the null hypothesis qualifies to be rejected, it indicates that data readings are strong and are probably not due to chance.


#### T-Test Assumptions

- The first assumption made regarding t-tests concerns the scale of measurement. The assumption for a t-test is that the scale of measurement applied to the data collected follows a continuous or ordinal scale, such as the scores for an IQ test.

- The second assumption made is that of a simple random sample, that the data is collected from a representative, randomly selected portion of the total population.

- The third assumption is the data, when plotted, results in a normal distribution, bell-shaped distribution curve.

- The final assumption is the homogeneity of variance. Homogeneous, or equal, variance exists when the standard deviations of samples are approximately equal.


#### Example

We are going to look at the income of the household for female/male headed household. The main_dataset doesn't contain the head of household gender information, so we are creating and randomly populating the gender variable.

```{r, warning=F}
set.seed(10)
main_dataset$hoh_gender = sample(c('Male', 'Female'), nrow(main_dataset), replace=TRUE)
main_dataset$b15_hohh_income = as.numeric(main_dataset$b15_hohh_income)

```

```{r}
t_test_results <- t.test(b15_hohh_income ~ hoh_gender, data = main_dataset)
t_test_results
```
In the result above :

t is the t-test statistic value (t = -0.19105),
df is the degrees of freedom (df= 1525.6),
p-value is the significance level of the t-test (p-value = 0.8485).
conf.int is the confidence interval of the means difference at 95% (conf.int = [-270.3420, 328.6883]);
sample estimates is the mean value of the sample (mean = 3459.044, 3429.871).

**Interpretation:**

The p-value of the test is 0.8485, which is higher than the significance level alpha = 0.05. We can't reject the null hypothesis and we can't conclude that men headed household average income is significantly different from women headed household average income. Which makes sense in this case given that the gender variable was randomly generated. 


### ANOVA

#### What is ANOVA 

Similar to T-test, ANOVA is a statistical test for estimating how a quantitative dependent variable changes according to the levels of one or more categorical independent variables. ANOVA tests whether there is a difference in means of the groups at each level of the independent variable.

The null hypothesis (H0) of the ANOVA is no difference in means, and the alternate hypothesis (Ha) is that the means are different from one another.

The T-test is used to compare the means between two groups, whereas ANOVA is used to compare the means among three or more groups.


#### Types of ANOVA

There are two main types of ANOVA: one-way (or unidirectional) and two-way.
A two-way ANOVA is an extension of the one-way ANOVA. With a one-way, you have one independent variable affecting a dependent variable. With a two-way ANOVA, there are two independents. For example, a two-way ANOVA allows a company to compare worker productivity based on two independent variables, such as salary and skill set. It is utilized to observe the interaction between the two factors and tests the effect of two factors at the same time.

#### Example

We are using the same example as the previous test.

```{r}

res.aov <- aov(b15_hohh_income ~ hoh_gender, data = main_dataset)

summary(res.aov)
```

As the p-value is more than the significance level 0.05, we can't conclude that there are significant differences between the hoh_gender groups.

#### Bonus : Visualization of the data to confirm the findinds 

```{r, warning=F}
library(ggpubr)

ggline(main_dataset, x = "hoh_gender", y = "b15_hohh_income", 
       add = c("mean_se", "jitter"), 
       order = c("Female", "Male"),
       ylab = "HH income", xlab = "hoh gender")
```


### chi-squares

Pearson`s chi-squared test.
Chi-squared test is a statistical hypothesis test, based on comparing expected and observed frequencies of 2-way distribution (2-way table).

There are three types of tests (hypothesis) that could be tested with chi-squared: independence of variables, homogenity of distribution of the characteristic across "stratas" and goodness of fit (whether the actual distribution is different from the hypothetical one). Two former versions use the same calculation.

Chi-squared test is a non-parametric test.
It is used with nominal to nominal (nominal to ordinal) scales.
For running chi-squared test, each cell of the crosstab should contain at least 5 observations.

Note, that chi-squared test indicates significance of differences in values on a *table level*. Thus, it is impossible to track significance in differences for particular options.

(For example, it is possible to say that satisfaction with public transport and living in urban / rural areas are connected. However, we can conclude from the test itself, that "satisfied" and "dissatisfied" options differ, while "indifferent" does not. By the way, it is possible to come up with this conclusion while interpreting a corresponding crosstab. Also, chi-squared test can tell nothing about the strength and the direction of a liaison between variables).


#### Chi-squared test for independence

The test for weighted data is being run with the survey library


```{r, message= F, warning=F, error=F, echo=T}
library(survey)
library(dplyr)
library(weights)
```

The use of the survey package functions requires specifying survey design object (which reslects sampling approach).

```{r}
main_dataset <- read.csv("inputs/UKR2007_MSNA20_HH_dataset_main_rcop.csv", na.strings = "", stringsAsFactors = F)

svy_design <- svydesign(data=main_dataset, 
                        id = ~1, 
                        strata = main_dataset$strata, 
                        probs = NULL,
                        weights = main_dataset$stratum.weight)

```

Specifying function for getting chi-squared outputs.

Hereby test Chi^2 test of independence is used. So, the test is performed to find out if strata variable and target variable are independent,
**H0** - the distribution of target variable values is equal across all stratas
**H1**  - strata and target var are not independent, so distribution varies across stratas.

Apart from getting test statistic and p value, we would like to detect manually what are the biggest and the smallest values across groups/stratas (in case if the difference is significant and it makes sense to do it).

Hereby test Chi^2 test of independence is used. So, the test is performed to find out if strata variable and target variable are independent,
H0 - the distribution of target var. values is equal across all stratas
H1  - strata and target var are not independent, so distribution varies across stratas.

Within the Survey package, the suggested function uses Scott & Rao adjustment for complex survey design

#####Simple single call of the test.

Let's say that the research aim is to understand whether the main water source is dependent on the area where HH lives (strata variable).

Thus, in our case:
H0 - the distribution of water source is equal across stratas and these variables are independent
H1 - the distribution of water source differs by strata and these variables are not independent.

Running the chi-squared test:

(Survey version - requires a survey design object)

```{r}
svychisq(~ f11_dinking_water_source + strata, design = svy_design, statistic = "Chisq")

```

The current output presents the result of Pearson's chi-squared test (Rao & Scott adjustment accounts for weighted data) [https://www.researchgate.net/publication/38360147_On_Simple_Adjustments_to_Chi-Square_Tests_with_Sample_Survey_Data].

The output consist of three parameters:
1. X-squared - observed value of chi-squared parameter,
2. df - "degrees of freedom" or number of independent categories in data -1
Having only these two elements it is possible to see, if with the current number of df, the value of chi-squared parameter is greater than the critical value for the chosen level of probability, using (tables of chi-squared critical values) [https://www.itl.nist.gov/div898/handbook/eda/section3/eda3674.htm]
Fortunately, R provides us with the third parameter, which allows us to skip this step.
3. p-value - stands for the probability of getting value of the chi-squared parameter greater or equal to the given one assuming that H0 is true. If p-value < 0.05 (for 95% confidence level, or 0.01 for 99%), the H0 can be rejected.

In the current example, as p-value is very low, H0 can be rejected, and hence, strata variable and drinking water source are not independent. 

(Weights package - can work with the dataframe directly, without the survey design object)

```{r}
wtd.chi.sq(var1 = main_dataset$f11_dinking_water_source, var2 = main_dataset$strata, weight = main_dataset$stratum.weight)
```


Specifying the function for crosstabs with chi-squared tests for independence.

#####Functions to have test results along with crosstabs

1. The function with survey design object. 

Apart from getting test statistic and p value, we would like to detect manually what are the biggest and the smallest values across groups/stratas (in case if the difference is significant and it makes sense to do it).

The function also helps to find out smallest and the largest values per row 

```{r}
chi2_ind_extr <- function(x, y, design){
  my_var <- y

# running Pearson's chi-squared test
  s_chi <- svychisq(as.formula(paste0("~", x,"+", y)), design, statistic = "Chisq")
  
  # extracting basic parameters
  wmy_t <- as.data.frame(s_chi$observed)
  wmy_t$p <- s_chi$p.value
  wmy_t$chi_adj <- s_chi$statistic
  wmy_t$df <- s_chi$parameter
  #indicating significance
  wmy_t$independence <- ifelse(wmy_t$p < 0.05, "h1", "h0")
  colnames(wmy_t)[1:2] <- c("var_x", "var_y")
  wmy_t_w <- spread(wmy_t, var_y, Freq)
  wmy_t_w[,1] <- paste0(x, " ", wmy_t_w[,1])
  colnames(wmy_t_w)[1] <- "variable"
  # getting percent from weighted frequencies
  cols_no_prop <- c("variable", "p", "independence", "max", "min", "chi_adj", "df")
  wmy_t_w[, !colnames(wmy_t_w) %in% cols_no_prop] <- lapply(wmy_t_w[, !colnames(wmy_t_w) %in% cols_no_prop], function(x){x/sum(x, na.rm = T)})
  # indicating extremum values
  cols_to_compare <- colnames(wmy_t_w)[!colnames(wmy_t_w) %in% c("p", "chi_adj", "df", "independence", "min", "max", "variable")]
  wmy_t_w$max <- ifelse(wmy_t_w$independence == "h1", cols_to_compare[apply(wmy_t_w, 1, which.max)], 0)
  
  wmy_t_w$min <- ifelse(wmy_t_w$independence == "h1", cols_to_compare[apply(wmy_t_w, 1, which.min)], 0)
  
  #adding Overall value to the table
  ovl_tab <- as.data.frame(prop.table(svytable(as.formula(paste0("~",x)), design)))
  colnames(ovl_tab) <- c("options", "overall")
  wmy_t_w <- wmy_t_w %>% bind_cols(ovl_tab)
  return(wmy_t_w)
}
```

The use with a single-choice question

```{r, warning=F}
chi2_ind_extr("f11_dinking_water_source", "strata", svy_design)
```
In the given output, **p** indicates chi-squared test p-value, **chi_adj** - value of chi-squared parameter, **df** - degrees of freedom, **independence** - indicates which hypothesis is accepted, h0 or h1. In the current function, H1 is accepted if p < 0.05. In case H0 is rejected and H1 accepted, in **min** and **max** columns minimum and maximum y variable categories are given.

The use with a multiple-choice question.

The test would not work if any of dummies consists from zeros only, it need to be checked in advance, thus, let's filter such variables out from the vector of variable names

Getting full range of dummies from the question

```{r}
names <- names(main_dataset)
f12 <- names[grepl("f12_drinking_water_treat.", names)]
```

Defining a function which would remove zero-sum variables for us

if the absence of empty cells for each strata is needed, per_cell = T
```{r}
non_zero_for_test <- function(var, strata, dataset, per_cell = F){
  sum <- as.data.frame(sapply(dataset[, names(dataset) %in% var], sum))
  sum <- cbind(rownames(sum), sum)
  filtered <- sum[sum[,2] > 0,1]
  tab <- as.data.frame(sapply(dataset[,names(dataset) %in% filtered], table, dataset[, strata]))
  ntab <- names(tab)
  #if the absence of empty cells for each strata is needed, per_cell = T
  tab <- tab[, which(colnames(tab) %in% ntab)]
  zeros <- apply(apply(tab, 2, function(x) x==0), 2, any)
  zeros <- zeros[zeros == F]
  zeros <- as.data.frame(zeros)
  zeros <- cbind(rownames(zeros), zeros)
  zeros_lab <- as.vector(unlist(strsplit(zeros[,1], split = ".Freq")))
  result <- c()
  if(per_cell == T){
    result <- zeros_lab
  } else {
    result <- ntab
  }
  message(paste0(abs(length(var) - length(result)), " variables were removed because of containing zeros in stratas"))
  
  return(result)
}
```

Using the test with multiple choice question.

```{r, warning=F}
nz_f12 <- non_zero_for_test(f12, "strata", main_dataset, per_cell = F)

f12_sig <- lapply(nz_f12, chi2_ind_extr, y = "strata", design = svy_design)

f12_sig_df <- Reduce(rbind, f12_sig)

#removing 0 ("not selected") options
f12_sig_df <- f12_sig_df %>% filter(options == 1)
f12_sig_df
```

2. The function without the survey design object

```{r}
chi2_ind_wtd <- function(x, y, weight, dataset){
  i <- 1
  table_new <- data.frame()
  for (i in 1:length(x)) {
  ni <- x[i]
  ti <- dataset %>% 
    filter(!is.na(!!sym(y))) %>%
    group_by(!!sym(y), !!sym(ni)) %>% 
    filter(!is.na(!!sym(ni))) %>%
    summarise(wtn = sum(!!sym(weight)), n = n()) %>%
    mutate(prop = wtn/sum(wtn), varnam = ni)
  names(ti) <- c("var_ind", "var_dep_val", "wtn", "n",  "prop", "dep_var_nam")
  ti <- ti %>% ungroup() %>% mutate(base = sum(c_across(n)))
  chi_test  <- wtd.chi.sq(var1 = dataset[, paste0(y)], var2 = dataset[, paste0(ni)], weight = dataset[, paste0(weight)])
  ti$chisq <- chi_test[1]
  ti$df <- chi_test[2]
  ti$p <- chi_test[3]
  ti$hypo <- ifelse(ti$p < 0.05, "h1", "h0")
  table_new <- rbind(table_new, ti)
}
table_cl_new_1 <- table_new %>% 
  #filter(var_dep_val != "0") %>% 
  filter(!is.na(var_dep_val))
#mutate(var_id = paste0(dep_var_nam, ".", var_dep_val))
table_cl_new_1 <- table_cl_new_1[, -c(3, 4)]

table_sprd <- tidyr::spread(table_cl_new_1, var_ind, prop) %>%
  arrange(dep_var_nam)
return(table_sprd)
}
```

The use with the single choice question

```{r}
chi2_ind_wtd(x = "f11_dinking_water_source", y = "strata",
             weight = "stratum.weight", main_dataset)
```
The use with multiple choice questions

```{r, message = F}
names <- names(main_dataset)

f12_names <- names[grepl("f12_drinking_water_treat.", names)]

f12_sig2 <- lapply(f12_names, chi2_ind_wtd, y = "strata", weight = "stratum.weight", main_dataset)

f12_tab_df <- Reduce(rbind, f12_sig2)
f12_tab_df
```

### Chi-squared Goodness of Fit test

Goodness of test chi-squared test is used to compare the observed distribution with the hypothetical one,

For instance, we would like to test the hypothesis, that 90% of HHs had soap in their HHs at the moment of the survey.

```{r}
main_dataset <- main_dataset %>% mutate(soap = case_when(
 p16_soap_household == "dont_know_refuse_to_answer" ~ "NA",
 p16_soap_household == "no" ~ "no",
 p16_soap_household == "yes_soap_is_not_shown" ~ "yes",
 p16_soap_household == "yes_soap_is_shown" ~ "yes"
))
soap_prob <- as.data.frame(prop.table(table(main_dataset$soap)))
soap_prob <- soap_prob[, 2]
# please, note that this is a function call with unweighted data!
chisq.test(soap_prob, p = c(0, 0.1, 0.9))
```
As p value is lower than the chosen confidence level, hypothesis of equality of the current distribution to hypothetical 90% / 10% of soap-owners can be rejected.

Defining a function for checking the distribution within a particular group, against the "hypothetical" overall sample.

For the overall sample we would need a variable which has the same value across all observations (e.g. informed consent == yes etc.)

```{r}

chi2_GoF <- function(name, stratum_hyp, stratum_2, value_stratum_hyp, value_stratum_2, data, weights){
  i <- 1
  table_1 <- data.frame()
  for (i in 1:length(name)) {
    ni <- name[i]
    ti <- data %>% 
      group_by(!!sym(ni)) %>% filter(!!sym(stratum_hyp) == paste0(value_stratum_hyp)) %>%
      filter(!is.na(!!sym(ni))) %>%
      summarise(wtn = sum(!!sym(weights)), n = n()) %>%
      mutate(prop = wtn/sum(wtn), base = sum(wtn), cnt = sum(n))%>%
      mutate(varnam = ni)
    names(ti) <- c("var", "wtn", "n", "prop", "base_w", "count", "nam")
    table_1 <- rbind(table_1, ti)
  }
  table_cl_1 <- table_1 %>% filter(!is.na(var)) %>%
    mutate(var_id = paste0(nam, ".", var))
  i <- 1
  table_2 <- data.frame()
  for (i in 1:length(name)) {
    ni <- name[i]
    ti <- data %>% 
      group_by(!!sym(ni)) %>% filter(!!sym(stratum_2) == paste0(value_stratum_2)) %>%
      filter(!is.na(!!sym(ni))) %>%
      summarise(wtn = sum(!!sym(weights)), n = n()) %>%
      mutate(prop = wtn/sum(wtn), base = sum(wtn), cnt = sum(n))%>%
      mutate(varnam = ni)
    names(ti) <- c("var", "wtn", "n", "prop", "base_w", "count", "nam")
    table_2 <- rbind(table_2, ti)
  }
  
  table_cl_2 <- table_2 %>% filter(!is.na(var)) %>%
    mutate(var_id = paste0(nam, ".", var))
  table_cl_compare <- merge(table_cl_1, table_cl_2, by = "var_id", all = T)
  names(table_cl_compare) <- c("var.id", "var.hyp", "wtn.hyp", "n.hyp", "prop.hyp", "base.hyp", "base_nw.hyp","nam.hyp", "var.2", "wtn.2", "n.2", "prop.2", "base.2", "base_nw.2", "nam.2")
  
  #Dealing with NA issues
  table_cl_compare$nam.hyp[is.na(table_cl_compare$nam.hyp)] <- table_cl_compare$nam.2[is.na(table_cl_compare$nam.hyp)]
  
  table_cl_compare$nam.2[is.na(table_cl_compare$nam.2)] <- table_cl_compare$nam.hyp[is.na(table_cl_compare$nam.2)]
  
  table_cl_compare$to_split <- table_cl_compare$nam.hyp
  
  table_cl_compare <- separate(table_cl_compare, col = to_split, into = c("indctr", "item"), sep ="\\.")
  
  table_cl_compare <- table_cl_compare %>% group_by(indctr) %>% fill(base_nw.hyp, .direction  = "down") %>% fill(base_nw.hyp, .direction  = "up") %>% ungroup()
  table_cl_compare <- table_cl_compare %>% group_by(indctr) %>% fill(base.hyp, .direction  = "down") %>% fill(base.hyp, .direction  = "up") %>% ungroup()
  table_cl_compare <- table_cl_compare %>% group_by(indctr) %>% fill(base_nw.2, .direction  = "down") %>% fill(base_nw.2, .direction  = "up") %>% ungroup()
  table_cl_compare <- table_cl_compare %>% group_by(indctr) %>% fill(base.2) %>% fill(base_nw.hyp, .direction  = "down") %>% fill(base.2, .direction  = "up") %>% ungroup()
  
  table_cl_compare$prop.hyp[is.na(table_cl_compare$prop.hyp)] <- 0
  table_cl_compare$prop.2[is.na(table_cl_compare$prop.2)] <- 0
  
  table_cl_compare$wtn.hyp[is.na(table_cl_compare$wtn.hyp)] <- 0
  table_cl_compare$wtn.2[is.na(table_cl_compare$wtn.2)] <- 0
  
  table_cl_compare$chi2 <- chisq.test(table_cl_compare$wtn.2, p = table_cl_compare$prop.hyp)$statistic
  table_cl_compare$p <- chisq.test(table_cl_compare$wtn.2, p = table_cl_compare$prop.hyp)$p.value
  table_cl_compare$df <- chisq.test(table_cl_compare$wtn.2, p = table_cl_compare$prop.hyp)$df
  
  table_cl_compare$H0 <- ifelse(table_cl_compare$p <= 0.05, "H1 - different", "H0 - same")
  return(table_cl_compare)
}

```

Using the function with a single choice question

```{r}
chi2_GoF(name = "f11_dinking_water_source", stratum_hyp = "a1_informed_consent", stratum_2 = "strata", value_stratum_hyp = "yes", value_stratum_2 = "5km_rural", data = main_dataset, weights = "stratum.weight")
```
Using the function with the multiple responce set

```{r, warning=F}
names <- names(main_dataset)

f12_names <- names[grepl("f12_drinking_water_treat.", names)]

#removing empty variables from the MR set
nz_f12 <- non_zero_for_test(f12_names, "strata", main_dataset, per_cell = F)

table(main_dataset[, nz_f12[4]], main_dataset[, "strata"])


f12_sig2 <- lapply(nz_f12, chi2_GoF, stratum_hyp = "a1_informed_consent", stratum_2 = "strata", value_stratum_hyp = "yes", value_stratum_2 = "5km_rural", data = main_dataset, weights = "stratum.weight")

f12_sig2_df <- Reduce(rbind, f12_sig2)
f12_sig2_df
```



